AiPoker - Implementation of a Poker Agent

### 2014-10-31 report/paper.tex
Much progress on report

%% indicate IEEE Member or Student Member in form indicated below
\author{John Holl\'en, Robin Berntsson, Simon Bergst\"om}
\authorfooter{
%% insert punctuation at end of each item
\item
 John Holl\'en, Robin Berntsson, Simon Bergstr\"om
}

%% Abstract section.
\abstract{ 
We will probably write this section the last thing we do. 
}
%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Monte Carlo Tree Search, Texas Hold'em}

@@

%% the only exception to this rule is the \firstsection command
\firstsection{Background}
\maketitle 
Texas Hold'em is a popular variant of poker, not just to play with friends but also online against other people for money. Poker belongs with the most difficult kind of card game to solve discretely since it is an stochastic game with imperfect information. The challenge of creating the best poker agent has attracted a lot of people and is a popular subject in AI development.

There are many different games today and the strategy to solve these games depends on the type of game. In game theory games can be classified due to the knowledge if the transition to a new state within the game depends on a stochastic process and if there is some hidden information. Poker belongs with the most difficult kind of card game to solve discretely since it is an stochastic game with imperfect information.
\\*\\*
\textbf{Perfect Information vs Imperfect Information}\\*
In perfect information games both players can observe the complete state of the game at any given time, this is the case in games such as Chess and Go. Contrary to this we have imperfect information games where each player is unable to observe the complete state, one example for this is poker were neither player can see the other players pocket cards.
\\*\\*
\textbf{Deterministic vs Stochastic}\\*
In deterministic games the next state of the game is uniquely determined by an action taken at the current state. In contrast, a stochastic game is a game where the player don't have control over the transition to the next state. For example in poker where the new cards is drawn from a scrambled deck.

@@

In general imperfect information games require more complex reasoning than perfect information games and deterministic games is easier than stochastic. This places Chess and go in the ?easier? category while poker is in the ?hardest? one.

The theory behind Monte Carlo search trees will be studied within this project and tested as a part of the logic that will determine the actions of the poker agent. Smaller algorithms to improve the logic for the AI will be tested and implemented in combination of the Monte Carlo Search Tree to create the best possible poker agent.

@@

The insight that random sampling can be used to solve problems that is too complex to solve analytically can be used to evaluate the best action in huge game trees. For a given node N we can estimate its value by running a number of simulations of how the game may end from that node. The result from those simulations can then be used to give an approximation for the value of that state and as N grows it will converge to the result from a Minimax search.

@@

\section{Theory}
\subsection{Texas Hold'em no Limit}
There are many forms of poker, with the most popular being the Texas Hold?em. Texas Hold?em can be played in two different styles, limited and no-limits. In limited poker the amount you bet or raise is fixed, and there is an upper amount of times that you can raise. However in no limited there is no restriction to the amount of money or times you can raise.

@@

\textbf{Pre Flop}\\*
In this stage the players gets their two cards from the dealer called the pocket cards and the player who is first are able to either bet, call if he has not paid the big blind, check if he has paid the big blind or fold. Since you can not form a five combination of cards yet only a speculation of how good your pocket cards is. If a player bets or raise all other players get the opportunity to either call,raise or fold max three times, after that the player who has not fold moves on to the flop.
\\*\\*
\textbf{Flop}\\*
The flop is when the dealer puts out the three first cards on the table who is visible for all players and makes it possible for all players to form their first five combination and calculate how good cards they have. Like in the pre flop phase all players will make moves in the same procedure to see which who are still left to the turn phase.
\\*\\*
\textbf{Turn}\\*
Now the dealer adds one card to the flop so its totally four visible card on the table. The players could then form a new five card combination to see how good their cards are. Same as previous steps there will be some betting before going on to the river.
\\*\\*
\textbf{River}\\*
The dealer add another card so there is five cards on the table for the player to make their best five combinations of cards with. This was the last card that will be distributed so after some more betting there is time for the showdown.
\\*\\*
\textbf{Showdown}\\*
In this phase the final betting is completed and the players who has not folded will now show their cards in order to decide who is the winner of this round. The winner will get the full pot and then a new round will start with the pre flop.
\\*\\*
\textbf{The different five combination of cards}\\*
There are nine different types of combination of cards and here they are listed in descending order[3]:
\begin{itemize}
  \item \textit{Straight Flush}: When you have a five cards in a consecutive order of value with the same suit. If more than one player have the Straight Flush the player with the highest values of their cards wins, and the best type is the Royal Straight Flush when you have Ace,King,Queen and Knight with the same suit.
  \item \textit{Four of a kind}: It is a hand that contains four of the same number, which means in all the different suits. If different players have the same type of hand the hand with the highest four of a kind wins.
  \item \textit{Full House}:   It is a hand where you have three cards of the same value in combination with two other cards that have the similar value. If more than one player have a Full House the player with the highest value three of a kind wins, if two players have the same three of a kind the player with the highest value of one pair wins.
  \item \textit{Flush}: It is when you have five cards with the same suit, if more than one player have a flush the one which have the highest card wins. If they have the same highest card then the the second highest card is checked and so on.
  \item \textit{Straight}: When you have a five cards in a consecutive order of value and if more than one player have a Straight then the player with the highest card wins.
  \item \textit{Three of a Kind}: When you have three cards of the same value, if more than one player has a Three of a kind then the player with the highest value of the cards wins.
  \item \textit{Two Pair}: When you have two card with the same values in combination with two other cards with the same values. In other words they have two pair of cards, if more than one player have a two pair the player with the highest pair has the best hand between them.
  \item \textit{One Pair}: When you have one pair of cards, if more than one player have a one pair the player with the highest pair has the best hand between them.
  \item \textit{High Card}: If the card does not match any of the types above then the card with the highest value is considered.

\end{itemize}

\subsection{Characteristics of a Poker Player}
Like previously described poker is a stochastic game, that makes it very hard to control and to define what a good poker player is. A good poker player does not always need to be the one who is playing optimally. A good poker player is hard to define but some of the main characteristics of a good player is that he is enjoying the game, can read his opponents strategy or way of playing and also know when having a good hand to play with.[REF!]

A player who tries to maximise his winnings will have to use bluffing. The reason for this is two-fold. First, since betting is often associated with holding strong cards, opponents might be tricked to give up a stronger hand. By betting on his weaker cards the bluffing player might win a pot that he would otherwise lose because he holds the weakest cards. Second and most interestingly: if a player is known to only bet with strong cards, his opponents can just fold whenever he bets and he will have a hard time winning money with his strong cards.

\section{Method}
\subsection{Basic Game Engine}
In order to even begin implementing an AI, the core poker game had to be implemented first. For simplicity the game implemented in this study is a two player game where a human player faces a computer AI. The player who starts is chosen at random when the game is started. The player that starts then puts one dollar in to the pot. The other player puts two dollars in the pot. Then the game is turn based like ordinary Texas Hold?em, the players can bet, check, call, raise and fold. And after each betting round cards are dealt to the table and the players combine the cards on hand with the cards on the table in order to get the strongest hand possible.

An very easy poker agent was implemented that played totally random was created in the purpose of testing the implemented game and later on the AI poker agent that was going to be implemented with Monte Carlo and other interesting poker algorithms(See 3.2.).

@@

\begin{equation} \label{eq:upperconfidence}
	v_{i} + C \cdot \sqrt{\frac{\ln N}{n_{i}}}
\end{equation}
The first term vi is the estimated value of the node and is responsible for exploitation of good nodes, the second term is made up by the total number of visits of the parent node $N$ and the total number of visits of the current node $n_{i}$. The constant $C$ allows tuning of the exploration-exploitation trade-off. 
\\*\\*
\textbf{Expansion}
\\*If the leaf node reached is not an terminating node, which means that the game is not over at the current node. Then create one or more child nodes from the node currently at. This step is illustrated in fig(\ref{fig:expansion}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree1.png}
    \caption{\label{fig:expansion} Illustration of how the green node is added to the tree.}
  \end{center}
\end{figure}
\\*
\textbf{Simulation}
\\*The third step in MCTS is the simulation step. Here a game is simulated from the state in the new node until a result is achieved. The value of the reached result is recorded.
\\*\\*
\textbf{Backpropagation}
\\*The last step is to update the expected value and the selection counter for all nodes along the explored path by backpropagating the recorded result from the simulation step. This step is illustrated in fig(\ref{fig:backprop}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree2.png}
    \caption{\label{fig:backprop} Illustration of how the path to the node is updated with the score from the simulation in the whole tree.}
  \end{center}
\end{figure}
When the back propagation step is finished the whole chain starts over again starting with the selection step. This is done until the either the game tree is fully searched or a time limit is reached. The fact that MCTS can give a result after a given time is a result of it being anytime, that is it will give a valid solution to the problem even if it is interrupted before it is finished. However the quality of the solution is an estimate and is expected to be better the more time the MCTS keeps running.

One problem that is introduced with the MCTS is how to run the simulation step. Here we have to select the moves for each player until the end of the game. One solution could be to just select random moves, this can work well for simple games but for games as complex as poker this will lead to bad results. This implies that we somehow have to model the opponent to be able to determine the correct actions in the simulation.

@@

@@

@@

\subsection{AI}
Since a number of algorithms were implemented in this study several results were achieved and therefore comparisons between the different approaches can be made. During this study, the different AI were put against each other playing several games. The following setups were made: 
\begin{itemize}
\item \textit{Monte Carlo + Hand Strength} vs \textit{Random Agent}
\item \textit{Effective Hand Strength} vs \textit{Monte Carlo + Hand Strength}
\item \textit{Monte Carlo} vs \textit{Monte Carlo + Hand Strength}
\item \textit{Monte Carlo + Hand Strength} vs \textit{Human Player}
\end{itemize}

\section{Conclusion}

@@

\bibitem{byron}
  Byron L., Wattenberg M.
  \emph{Stacked graphs - Geometry \& Aesthetics}.
  IEEE Transactions on Visualization and Computer Graphics archive
Volume 14 Issue 6,
Pages 1245-1252. IEEE Educational Activities Department Piscataway, NJ, USA
ISSN: 1077-2626.
  2008.

\bibitem{reynolds}
  Reynolds A. P., Richards G., Rayward-Smith V. J.
  \emph{The Application of K-medoids and PAM to the Clustering of Rules}.
  In: Intelligent Data Engineering and Automated Learning - IDEAL 2004. Lecture Notes in Computer Science, 3177 . Springer-Verlag, pp. 173-178. ISBN 978-3-540-22881-3,
  2004.

>>>>

%% indicate IEEE Member or Student Member in form indicated below
\author{John Holl\'en, Robin Berntsson, Simon Bergstr\"om}
\authorfooter{
%% insert punctuation at end of each item
\item
 John Holl\'en, Robin Berntsson, Simon Bergstr\"om
}

%% Abstract section.
\abstract{ 
Poker is a well known card game worldwide and has become very popular to play online for the past few years. Together with the interest of playing the game online an interest of trying to create the best poker agent for online poker became popular. This report describes a project to investigate the possibilities with creating a Poker Agent with the logic of a Monte Carlo Tree Search in combination a few smaller well known algorithms. The result is a poker agent which plays with realistic reasoning.
}
%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Monte Carlo Tree Search, Texas Hold'em}

@@

%% the only exception to this rule is the \firstsection command
\firstsection{Background}
\maketitle 
Texas Hold?em is a popular variant of poker, not just to play with friends but also online against other people for money. The challenge of creating the best poker agent has attracted a lot of people and is a popular subject in AI development.
\\*\\*
\textbf{Type of Games}\\*
There are many different games today and the strategy to solve these games depends on the type of game. In game theory games can be classified due to the knowledge if the transition to a new state within the game depends on a stochastic process and if there is some hidden information. Poker belongs with the most difficult kind of card game to solve discretely since it is an stochastic game with imperfect information.
\\*\\*
\textbf{Perfect Information vs Imperfect Information}\\*
In perfect information games both players can observe the complete state of the game at any given time, this is the case in games such as Chess and Go. Contrary to this we have imperfect information games where each player is unable to observe the complete state, one example for this is poker were neither player can see the other players pocket cards. 
\\*\\*
\textbf{Deterministic vs Stochastic}\\*
In deterministic games the next state of the game is uniquely determined by an action taken at the current state. In contrast, a stochastic game is a game where the player don't have control over the transition to the next state. For example in poker where the new cards is drawn from a scrambled deck.

@@

The subdivision in (table \ref{tab:table1}) exhibit that games like Chess and Go belongs to the category of games that is less complex to try to solve with an AI agent. Poker belongs to one of the most complex games to try to solve.

\subsection{Related Work}
\subsubsection{Game Theory}
In 1997 Koller and Pfeffer made a program called Gala, the program uses strategies from game theory to find optimal strategies in imperfect information games. With the help of Gala they were able to solve several games but they come to the conclusion that the number of states in poker is too large to be solvable with the current hardware. However they suggested a solution that maybe it was possible to map several similar states onto one abstract state resulting in an abstract game tree much smaller than the original. 
It was from this idea that (Billings et al 2003) successfully abstracted the game down from a search space of $O(10^{18})$ to $O(10^{7})$. They then used Linear optimization to find a Nash equilibrium solution to the game. From this they could take any action, map it to the abstract game tree, find a solution and then map it back to the real game. This resulted in a poker agent (PsOPti) that was far better than anything else at the time.

If PsOpti would play a real nash equilibrium strategy no human player would be able to beat it in the long run. But the problem is that it is just an approximation and there will therefore find ways to exploit it. And as PsOpti ignore any action from the human player it allows the player to systematically search for weaknesses in the without fear of being punished.

\subsubsection{Expert Systems}
Billings et al. (1998) developed an AI agent called Loki. It is an expert system that uses a set of rules and formulas based on the strength of the cards to determine the next action. It uses no opponent modeling in heads up poker. It can be challenging to define strong expert systems as it can be hard to define rules that covers all aspects of the game. There can also be problems with conflicting rules and missing rules (situations that is not covered by any rules). Another problem is that expert systems require that you have great knowledge about the game as the AI will never play better then the player defining the rules.

\subsubsection{Opponent Modeling}
Papp (1998) introduced the first tries to extend Loki with opponent modeling. The opponent model was a quite simple and just counted the opponents actions in the different streets (preflop, flop, turn, river), but even this simple opponent modeling increased the strength of the AI.

Davidson et al. (2000) tried to improve the opponent modeling in Loki with artificial neural networks (ANNs) to predict the opponent?s actions, and achieves about 85\% accuracy at predicting the moves. This is a significant improvement over the previous implementation and the new system was named Poki. The problems with ANNs is that they can be slow to train and thus they will make weak plays against players who frequently change their play style.

\subsubsection{Search Trees}
Billings et al. (2004) was one of the first to use game trees search in a poker AI. Their Ai, Vexbot uses two modified versions of Expectiminimax called Miximax and Miximix that extends Expectiminimax to games with hidden information. Vexbot makes an exhaustive  depth first search of the entire search tree for two player poker.

Van den Broeck (2009) and Van den Broeck et al. (2009) were the first to apply Monte Carlo search to game trees for poker agents. They focused on 10 player no-limit poker which have enormous game trees which makes methods such as the one described in Billings et al. (2004) impossible to use. Here it is impossible to search the whole game tree which makes sampling parts of the game tree the only option.

@@

The insight that random sampling can be used to solve problems that is too complex to solve analytically can be used to evaluate the best action in huge game trees. For a given node $N$ we can estimate its value by running a number of simulations of how the game may end from that node. The result from those simulations can then be used to give an approximation for the value of that state and as $N$ grows it will converge to the result from a Minimax search.

@@

\section{Theory}
\subsection{Texas Hold'em no Limit}
\textit{(If you already are familiar with the game of Texas Hold'em, you can skip to section 2.2).}
There are many forms of poker, with the most popular being the Texas Hold'em. Texas Hold?em can be played in two different styles, limited and no-limits. In limited poker the amount you bet or raise is fixed, and there is an upper amount of times that you can raise. However in no limited there is no restriction to the amount of money or times you can raise.

@@

\textbf{Pre Flop}\\*
In this stage the players gets their two cards from the dealer called the pocket cards and the player who is first are able to either bet, call if he has not paid the big blind, check if he has paid the big blind or fold. Since you can not form a five combination of cards yet only a speculation of how good your pocket cards are can be made. If a player bets or raises all other players get the opportunity to either call, raise or fold max three times, after that the players who has not folded moves on to the flop.
\\*\\*
\textbf{Flop}\\*
The flop is when the dealer puts out the three first cards on the table. These cards are visible to all players. The pocket cards can now be combined with the flop cards in order to create a strong hand. The players will make moves in the same procedure as in the pre flop stage.
\\*\\*
\textbf{Turn}\\*
Now the dealer adds a fourth card to the flop. The players will now bet again.
\\*\\*
\textbf{River}\\*
The dealer adds a fifth card to the table. This is the last card which is dealt. After the last betting round, the game moves on to showdown. 
\\*\\*
\textbf{Showdown}\\*
In this phase the final betting is completed and the players who has not folded will now show their cards in order to decide who is the winner of this round. The winner will get the full pot and then a new round will start with the pre flop.
\\*\\*
\textbf{The different five combination of cards}\\*
There are nine different types of combination of cards and here they are listed in descending order[3]:
\begin{itemize}
  \item \textit{Straight Flush}: When the player has five cards in a consecutive order of value with the same suit. If more than one player have a Straight Flush the player with the highest values of his/her cards wins. The best variant is the Royal Straight Flush when the player has Ace, King, Queen, Knight and a ten within the same suit.
  \item \textit{Four of a kind}: It is a hand that contains four of the same number, which means in all the different suits. If different players have the same type of hand the hand with the highest four of a kind wins.
  \item \textit{Full House}: This hand is a combination of three of a kind and a pair. If more than one player have a Full House the player with the highest value three of a kind wins, if two players have the same three of a kind the player with the highest value of one pair wins.
  \item \textit{Flush}: A flush is when a player has five cards with the same suit, if more than one player have a flush the one with the highest card wins. If they have the same highest card then the the second highest card is checked and so on.
  \item \textit{Straight}: When you have a five cards in a consecutive order of value. If more than one player has a Straight then the player with the highest card wins.
  \item \textit{Three of a Kind}: When you have three cards of the same value, if more than one player has a Three of a kind then the player with the highest value of the cards wins.
  \item \textit{Two Pair}: When a player has two pair of cards, if more than one player have a two pair the player with the highest pair has the best hand between them.
  \item \textit{One Pair}: When you have one pair of cards, if more than one player have a one pair the player with the highest pair has the best hand between them.
  \item \textit{High Card}: If the card does not match any of the types above then the card with the highest value is considered. 
\end{itemize}

These rules are the most basic rules and some special cases exists that is not described above, for a more detailed guide of the rules visit : http://www.pokerstars.se/poker/games/rules/hand-rankings/ and http://www.pokerstars.se/poker/games/rules/

\subsection{Characteristics of a Poker Player}
Like previously described poker is a stochastic game, that makes it hard to control and to define what a good poker player is. A good poker player does not always have to be the one who is playing optimally. A good poker player is hard to define but some of the main characteristics of a good player is that he can read his opponents strategy or way of playing, know when having a good hand to play with and know when to play the hand.

A player who tries to maximise his winnings will have to use bluffing. This applies both to the cases when the player holds good cards, and when he holds bad cards. If a player never bluffs his or her strategy becomes predictable and it becomes obvious that the player has good cards as soon as he or she bets. This will lead to the opponent folding and the size of the winning pot being small. 
In the second case, the player might have bad cards. If the player uses bluffing it is possible for him/her to win games where the opponent has better cards.

\subsection{Game Trees and the Motivation for Monte-Carlo Tree Search}
A game tree is a tree structure that from a given state in the game describes all moves to all possible endings of the game.
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.30]{img/MiniMaxGameTree.png}
    \caption{\label{fig:gametree} Part of a tic-tac-toe - game tree.}
  \end{center}
\end{figure}

The above example shows a part of the game tree for a tic-tac-toe game tree. Every node represent a possible state in the game with the root node being the current state and every edge is the action leading to that state. 
The size for a game tree grows very fast when you increase the possible actions that can be taken in each state. For example the game tree for a empty tic-tac-toe board have $255168$ leaf nodes. Two player Texas Hold'em have a search space size of $O(10^{18})$.

With the help of game trees computers can reason about the consequences of their actions and many of the most used AI algorithms are based on them. One of the most used tree search algorithms is the Minimax algorithm. Minimax assumes that both players play an optimal strategy. This means that the player who tries to find the best move always selects the actions that will give him the highest reward thus we call his nodes MAX-nodes. On the other hand the opponent will always select the actions that will give the player the lowest reward and is thus called his nodes MIN-nodes. The value for each leaf node is determined by a heuristic and the value is back propagated up the tree depending on the parent node. Minimax will search the game tree depth first and the result will be the action which gives the highest reward.

Expectiminimax is the continuation of the Minimax algorithm that allows for stochastic events in the game tree such as drawing cards. Expectiminimax trees add an additional type of nodes, chance nodes. At chance nodes a child $c$ is chosen with some probability $P(c)$. This means that a chance node for drawing a card from a full deck have $52$ children with a uniform probability distribution of $1/52$. The value for a chance node is the weighted average of all its children.

The problem with Minimax and Expectiminimax is that they use depth first search, this poses a problem for games such as poker which have such a big game tree that it can not be fully searched in a reasonable amount of time.

It was this problem that led to the development of Monte-Carlo tree search(MCTS). Monte-Carlo tree search works around the need to search the game tree exhaustively by estimating the best action by sampling the tree in its most promising parts.

\section{Method}
\subsection{Basic Game Engine}
In order to even begin implementing an AI, the core poker game had to be implemented first. This was made by first implementing an agent that played at random. For simplicity the game implemented in this study is a two player game with the limited edition of Texas Hold?em where a human player faces a computer AI. The player who starts is chosen at random when the game is started. The player that starts then puts one dollar in to the pot. The other player puts two dollars in the pot. Then the game is turn based like ordinary Texas Hold?em, the players can bet, check, call, raise and fold. And after each betting round cards are dealt to the table and the players combine the cards on hand with the cards on the table in order to get the strongest hand possible.

@@

\begin{equation} \label{eq:upperconfidence}
	v_{i} + C \cdot \sqrt{\frac{\ln N}{n_{i}}}
\end{equation}
The first term vi is the estimated value of the node and is responsible for exploitation of good nodes, the second term is made up by the total number of visits of the parent node $N$ and the total number of visits of the current node $n_{i}$. The constant $C$ allows tuning of the exploration-exploitation trade-off. 
\\*\\*
\textbf{Expansion}
\\*If the leaf node reached is not an terminating node, which means that the game is not over at the current node. Then create one or more child nodes from the node currently at. This step is illustrated in fig(\ref{fig:expansion}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree1.png}
    \caption{\label{fig:expansion} Illustration of how the green node is added to the tree.}
  \end{center}
\end{figure}
\\*
\textbf{Simulation}
\\*The third step in MCTS is the simulation step. Here a game is simulated from the state in the new node until a result is achieved. The value of the reached result is recorded.
\\*\\*
\textbf{Backpropagation}
\\*The last step is to update the expected value and the selection counter for all nodes along the explored path by backpropagating the recorded result from the simulation step. This step is illustrated in fig(\ref{fig:backprop}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree2.png}
    \caption{\label{fig:backprop} Illustration of how the path to the node is updated with the score from the simulation in the whole tree.}
  \end{center}
\end{figure}
When the backpropagation has been done the whole chain starts over again starting with the selection stage. [REF!] The use of the UTC selection method makes the game tree grow asymmetrically as it focus its search in the more promising parts of the tree.

One of the pros with MCTS is that it is anytime, that is it will give a valid solution to the problem even if it is interrupted before it is finished. However the quality of the solution is an estimate and is expected to be better the more time the MCTS keeps running.

@@

\subsubsection{Bucketing}
Since EHS needs at least the flop in order to be used for decision making, the Bucketing algorithm was implemented  as a complement to the EHS algorithm. This algorithm is only used in in the pre-flop phase where no board cards are visible.
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.60]{img/bucketlist.png}
    \caption{\label{fig:bucketlist} The bucket list.}
  \end{center}
\end{figure}

Fig(\ref{fig:bucketlist}) gives every two card combination a value between 0 and 4 where 4 is a good combination and 0 is a bad combination. This list is the source of decision making in the pre-flop phase when using the EHS-algorithm as logic in the poker agent.

\subsubsection{Randomness}
All previously mentioned algorithms take cards as input and gives the probability of the cards in different way to make the best decision with the current hand cards. Since a good poker player needs to be able to bluff, a random function has been implemented in order to give the agent a certain level of unpredictableness. The agent will make a random move approximately 5\% of the time.

@@

\subsection{AI}
The AI-file contains all methods necessary for the AI to be able to make a decision about what move to make. The method which is used the most is: findBestMove() which initializes a new instance of the SearchTree class and then runs a simulation from the current game state. It also contains methods for hand strength, hand potential and effective hand strength. A so called bucket is also present. Bucketing, which has been mentioned earlier is also present and is used as a heuristic in the first round before the flop has been presented on the table.

@@

\subsection{AI}
Since a number of algorithms were implemented in this study several results were achieved and therefore comparisons between the different approaches can be made. During this study, the different AI were put against each other playing several games. We have two different kinds of diagrams, one show the percentage of won games and one show the percentage of won hands. The following setups were made: 
\begin{itemize}
\item \textit{Monte Carlo + Hand Strength} vs \textit{Random Agent}
\item \textit{Effective Hand Strength} vs \textit{Monte Carlo + Hand Strength}
\item \textit{Monte Carlo} vs \textit{Monte Carlo + Hand Strength}
\item \textit{Monte Carlo + Hand Strength} vs \textit{Human Player}
\end{itemize}

\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.40]{img/randomvsmontehs2.png}
    \caption{\label{fig:randomvsmontehs} Result from Random vs Monte Carlo + Hand Strength.}
  \end{center}
\end{figure}

\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.40]{img/montehsvsehs.png}
    \caption{\label{fig:montehsvsehs} Result from Monte Carlo + Hand Strength vs EHS.}
  \end{center}
\end{figure}

\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.40]{img/montevsmontehsfixed.png}
    \caption{\label{fig:montevsmontehsfixed} Result from Monte Carlo vs Monte Carlo + HS (fixed number of simulations).}
  \end{center}
\end{figure}

\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.40]{img/montevsmontehs5sec.png}
    \caption{\label{fig:montevsmontehs5sec} Result from Monte Carlo vs Monte Carlo + HS (5sec time limit).}
  \end{center}
\end{figure}

\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.40]{img/montevsehs5sec.png}
    \caption{\label{fig:montevsehs5sec} Result from Monte Carlo vs EHS (5sec time limit).}
  \end{center}
\end{figure}

\subsubsection{\textit{Monte Carlo + Hand Strength} vs \textit{Human Player}}
 When we let humans play against the AI it achieved about 45\% win rate over all games. All humans that played against it rated themselves as relative weak poker players and it would probably achieve worse result versus more experienced players. Three different persons participated in the tests against the AI.

\section{Conclusion}
\subsection{Evaluation of the Result}
At a first glance it can look strange that the random ai have so many won hands, but one have to take into consideration that the amount of won hands is not the same as the amount of money won. The random agent wins a lot of hands, but this is due to the other agents folding when they have bad cards. So the average size of the money pot is very small for the hands the random bot wins and much larger for the hands the other agents win.

From the result it is clear that the Monte Carlo methods is clearly stronger players than the rule based expert system that uses effective hand strength (EHS) to choose its moves. The difference between the agents could possible be less if we spent more time refining the rules used by the EHS agent.

Another interesting observation is that the Monte Carlo agent that uses the hand strength to determine the moves in the simulation step loses to the one who does not use it if both is constrained to 5 sec of simulations. The reason for this is that calculating the hand strength takes so much time that the agent who does not use it can run 200 times more simulations. This shows that increasing the quality of the simulations is not always justified.

On the other hand if we let both simulations run a fixed amount of simulations we get 100\% win rate for the Monte Carlo agent that uses hand strength.

\subsection{Future Work}
\subsubsection{Opponent Modeling}
The quality of the solution given a certain number of iterations depends greatly on how good the simulation step can model the reality. For simple games it may be enough to simply use random moves but for games as complicated as poker this will lead to a weak agent. The reason for this is that a player's actions is highly dependent on what cards he holds. Note that we are not interested in finding the most probable move of the opponent but the probability distribution over the available moves. The reason for this is that if we only get the most probable move then each simulation will result in the same sequence of actions.

One simple approach is to model each opponent as a straight forward player, that means that we assume that an aggressive move (raise, bet) indicates a strong hand, and a passive move (check, call) indicates a weak hand. This can be  implemented by holding a histogram with all streets (pre-flop, flop, turn, river) and all the actions taken by the player in those states (fold, bet, call, raise, check). A raise in a game state indicates that a player have a strong hand if he usually does not raise in that state and vice versa. If the game goes to showdown it is possible to calculate the hand strength of the opponents cards on hand and thus make it possible to associate a hand strength with the actions taken that round.

\subsubsection{Speed}
The quality of the MCTS solution depends on the amount of simulations it can make, the code in the project is far from optimized and it is probably easy to increase the number of simulations per round by refactoring and optimizing parts of the code.

\subsubsection{Evaluation Against Human Player}
All evaluation has been performed between two poker agents in order to see which algorithm is winning but no exhaustive tests has been performed against human player. Since human players with different level of skills in poker can better evaluate the poker agent and interpret of the agents move can easily be predicted or not.

\subsubsection{Add More Parameters in Move Decision}
During the research phase of this project, many different algorithms to calculate the best move were found. The algorithm used in this project  was chosen by the reason of time range for the project and the estimated popularity of the algorithm. The different algorithms usually evaluate different characteristics of the current state to choose a move. This enabled different algorithms to be combined.

@@

\bibitem{levente}
  Levente Kocsis and Csaba Szepesvári. 2006. Bandit based monte-carlo planning. In Proceedings of the 17th European conference on Machine Learning (ECML'06), Johannes Fürnkranz, Tobias Scheffer, and Myra Spiliopoulou (Eds.). Springer-Verlag, Berlin, Heidelberg, 282-293. DOI=10.1007/11871842\_29

\bibitem{grze}
Grzegorz Fedczyszyn, Leszek Koszalka, and Iwona Pozniak-Koszalka. 2012. Opponent modeling in texas hold'em poker. In Proceedings of the 4th international conference on Computational Collective Intelligence: technologies and applications - Volume Part II (ICCCI'12), Ngoc-Thanh Nguyen, Kiem Hoang, and Piotr J?drzejowicz (Eds.), Vol. Part II. Springer-Verlag, Berlin, Heidelberg, 182-191. DOI=10.1007/978-3-642-34707-8\_19

\bibitem{billingsburch}
 D. Billings, N. Burch, A. Davidson, R. Holte, J. Schaeffer, T. Schauenberg, and D. Szafron. 2003. Approximating game-theoretic optimal strategies for full-scale poker. In Proceedings of the 18th international joint conference on Artificial intelligence (IJCAI'03). Morgan Kaufmann Publishers Inc., San Francisco, CA, USA, 661-668.

\bibitem{billingspapp}
D. Billings, D. Papp, J. Schaeffer, and D. Szafron. Poker as a testbed for machine intelligence research. In Proceedings of Advances in Artificial Intelligence Research, pages 1-15, 1998b.

\bibitem{papp}
Denis Richard Papp. 1998. Dealing with Imperfect Information in Poker. MS Thesis. University of Alberta, Edmonton, Alta., Canada. UML Order No. GAXMQ--34401.

\bibitem{davidson}
A. Davidson, D. Billings, J. Schaeffer, and D. Szafron. Improved opponent modeling in Poker. In Proceedings of the 2000 International Conference on Artificial Intelligence, pages 1467-1473, 2000.

\bibitem{broeck}
Guy Broeck, Kurt Driessens, and Jan Ramon. 2009. Monte-Carlo Tree Search in Poker Using Expected Reward Distributions. In Proceedings of the 1st Asian Conference on Machine Learning: Advances in Machine Learning (ACML '09), Zhi-Hua Zhou and Takashi Washio (Eds.). Springer-Verlag, Berlin, Heidelberg, 367-381. DOI=10.1007/978-3-642-05224-8\_28

\bibitem{billingsdavidson}
Darse Billings, Aaron Davidson, Terence Schauenberg, Neil Burch, Michael Bowling, Robert Holte, Jonathan Schaeffer, and Duane Szafron. 2004. Game-Tree search with adaptation in stochastic imperfect-information games. In Proceedings of the 4th international conference on Computers and Games (CG'04), H. Jaap Herik, Yngvi Björnsson, and Nathan S. Netanyahu (Eds.). Springer-Verlag, Berlin, Heidelberg, 21-34. DOI=10.1007/11674399\_2

\bibitem{schweizer}
Immanuel Schweizer, Kamill Panitzek, Sang-Hyeun Park, and Johannes Fürnkranz. 2009. An exploitative Monte-Carlo poker agent. In Proceedings of the 32nd annual German conference on Advances in artificial intelligence (KI'09), Bärbel Mertsching, Marcus Hund, and Zaheer Aziz (Eds.). Springer-Verlag, Berlin, Heidelberg, 65-72.

\bibitem{wenkai}
Wenkai Li and Lin Shang, "Estimating Winning Probability for Texas Hold'em Poker," International Journal of Machine Learning and Computing vol. 3, no. 1, pp. 70-74, 2013

\bibitem{pokerstars}
Rules for Texas Hold'em.(2014, October 31). Retrieved from http://www.pokerstars.se/poker/games/rules/

\bibitem{pokerlistings}
"Good Players vs. Winning Players in Poker" by Arthur Reber.(2011 January 6). Retrieved from http://www.pokerlistings.com/poker-strategy/good-players-vs-winning-players
